CENSE - Perceptual test II

### 2019-02-25 report.tex
proof read done

In the previous section we identified indicators based on masking models that can be computed from separated sources audio in the scene. These indicators provide estimates of the perceived time of presence that can be used to approximate pleasantness using a simple linear model with relatively low error. The objective of this section is to propose a deep learning architecture to predict the source-specific quantities needed for this approximation from mixed sound scenes.

@@

Based on the results outlined by the perceptual experiment, a processing chain is obtained for the estimation of pleasantness in urban sound scenes as shown in Figure~\ref{fig:ppred}. Pleasantness is modeled as a function of overall loudness and the perceived times of presence of traffic, voices and birds sources. However, predicting these quantities directly using deep learning architectures is difficult: training this type of models requires large annotated datasets which would require extensive work by several humans to obtain relevant labels. However, the perceived time of presence can be approximated by a simple masking model as demonstrated in Section~\ref{sec:physm}. While the same need for annotation of large datasets exists, it can be circumvented in the case of simulated data, for which the contribution of each source in the mix is available. Thus, a deep learning model can be used to predict either of the first steps of the pleasantness estimation chain.\\

An advantage of predicting sound levels is the accuracy of the labels as they are computed directly on available separated sources audio. On the condition that isolated samples used during scene generation are of good quality in terms of low other source content, the resulting ground truth labels can be considered exact. However, the prediction is a regression task which is typically difficult to learn. The best studied masking model relies on third-octave band sound levels, meaning that the learned model needs to perform an implicit source separation. Moreover, once an estimations of the sound levels are obtained the selected masking model must be used as the next step of pleasantness prediction. This model is not fully accurate and thus leads to small errors in pleasantness values even for perfect sound level estimations.\\

Instead, we can choose to directly predict if each of the three sources is present (heard) at a given time. This is either a detection problem if the data is processed asynchronously or a multi-label classification problem if the data is processed in frames, both of which are much simpler tasks. The corresponding labels are directly the outputs of the best studied masking model $\hat T(\alpha, \beta)$. Its uncertainties are thus transferred to "weak" labels rather than present as part of the constant processing chain used to obtain pleasantness estimations.\\

In this report two models for source-specific sound level and presence estimation are implemented and compared.

@@

Two datasets of simulated scenes are created for the proposed task:
\begin{itemize}
\item The development set containing 400 45~s scenes (total 5 hours) is used during the training process
\item The evaluation set containing 200 45~s scenes (total 2.5 hours) is used to compute the model's performance and generalisation capabilities
\end{itemize}

The generation procedure is the same for both datasets, and is equivalent to that of the simulated sub-corpus used during the perceptual test. Only traffic, human voices and birds sources compose the sound scenes due to the low number of available samples for other sources. The isolated samples database is split in the same proportions as the two datasets: two-thirds for the development set and the remaining one-third for the evaluation set. As a result, while the generation process is the same for both corpora both the scenarios and samples are different.\\

The sensors implemented as part of the Cense project provide fast measurements of sound levels for each third-octave band in the $20~Hz - 12.5~kHz$ range in their normal operation. Thus, the model should take as input the same representation. Third-octave spectra are computed for 125~ms frames and 29 frequency bands of the mixed sound scenes. Inputs are then obtained by splitting the resulting spectrograms into frames with a 1~s duration. This scale is considered relevant for perception and was already used to compute the masking models proposed in the perceptual experiment results analysis. The input frames of dimension 29x8 are processed independantly in this study.\\

@@

Training the models is achieved by progressively minimizing a loss function $L(y, \hat y)$, xhere $y$ and $\hat y$ represent the target and predicted outputs respectively. For sound level estimation we use the MSE or L2 loss:
\begin{equation}
L(y, \hat y) = \sum_s \left(\hat y_s - y_s\right)^2
\end{equation}
where $s$ is the source, $y_s$ and $\hat y_s$ are the target and predicted sound level of source $s$ in dB. For presence prediction the cost function is the binary cross-entropy of equation:
\begin{equation}
L(y, \hat y) = -\sum_s y_s log\left(\hat y_s\right) + (1-y_s) log\left(1-\hat y_s\right)
\end{equation}
where $s$ is the source, $y_s$ and $\hat y_s$ are the target and predicted presence for source $s$ in the 0-1 range.\\

@@

Table~\ref{tab:perf_cmp} displays the performance of the two studied models for metrics at the physical level. The first model outputs very poor estimations of the source-specific sound level, but the resulting source presence accuracy is high at 92.51\%. As a result, the time of presence estimates are precise with a root mean-squared error of 0.12 on a 0-1 scale. This behaviour may be explained by the permissiveness of the $\hat T_s(\alpha)$ indicator: if sources are close in terms of full-band sound level, the threshold $\alpha = -19dB$ allows rougher estimates for binary presence values. Thus, when using $\hat T_s(\alpha)$ it is sufficient to obtain good presence detection that the model output either very low values when the source is absent or poor estimates of the level when it is present. Here, the model fails at the source separation and level estimation task and is probably closer to a detection model.\\

The presence detection model performs similarly well for both presence accuracy (92.11\%) and resulting $\hat T_s(\alpha, \beta)$ estimation (0.12 RMSE). Both models obtain lower accuracies for birds than for traffic or voices. However, the origin of errors varies. The sound level model has an extremely low false positive rate consistently across sources while its false negative rate varies from 7.8\% for traffic to more than 18\% for birds. The reverse is visible for the presence model which displays a higher false positive rate for birds. Overall, traffic is found to be the most easily detected source, which may be due to its location in the lower frequency bands while voice and bird information is usually concentrated in narrower frequency ranges when using third-octave bands.\\

@@

Finally, a comparison is made of the models's capacity to result in relevant pleasantness estimates. For this analysis the perceptual experiment is used as assessments are available. It is further split in two parts:
\begin{enumerate}
\item The 25 recorded and replicated scenes, which contain additional sources not present in the development or evaluation datasets simulated for the optimization of deep learning models,
\item The 75 simulated scenes that are obtained from the same simulation process as both the development and evaluation dataset.
\end{enumerate}
The results are shown in Table~\ref{tab:pppred}. Again, both models lead to similarly correct estimations of pleasantness with errors lower than 1.20 on a 10-point scale. Interestingly, predictions are better on the first tested sub-corpus which contains other sources, indicating that models are quite robust to this type of perturbation in the data. Though, a larger set of scenes and environments with perceptual annotations would be required to further validate this point.\\

@@

Subsequently, a deep learning model with simple, state-of-the-art architecture is developed for this task. The model takes as input the fast third-octave band representation of the mixed scene and is trained on a corpus of simulated scenes to output presence labels in short time frames of 1s. Low pleasantness prediction errors are achieved on the perceptual experiment corpus: around 1.15 root mean-squared error on a 10-point scale.\\

Future work will be mainly focused on the study of the robustness of developed models to new environment and data. Here only three sources were considered during training, and a limited isolated sound sources database. Thus, pleasantness estimation chain needs to be adapted to the city of Lorient, in which the Cense sensor network is to be implemented, as it contains specific sound environments as well as unique sources that were not considered in this study.

>>>>

In the previous section,  indicators based on masking models that can be computed from separated sources audio in the scene are proposed. These indicators provide estimates of the perceived time of presence that can be used to approximate pleasantness using a simple linear model with relatively low prediction error. The objective of this section is to propose a deep learning architecture to predict the source-specific quantities needed for this approximation from mixed sound scenes.

@@

Based on the results outlined by the perceptual experiment, a processing chain is obtained for the estimation of pleasantness in urban sound scenes as shown in Figure~\ref{fig:ppred}. Pleasantness is modeled as a function of overall loudness and the perceived times of presence of traffic, voices and birds sources.

Predicting these quantities directly \ml{from what ?} using deep learning architectures is difficult: training this type of models requires large annotated datasets which would require extensive work by several humans to obtain relevant labels. However, the perceived time of presence can be approximated by a simple masking model as demonstrated in Section~\ref{sec:physm}. While the same need for annotation of large datasets exists, it can be circumvented by the use of simulated data, for which the contribution of each source in the mix is available. Thus, a deep learning model can be used to predict either of the first steps of the pleasantness estimation chain.\\

An advantage of predicting sound levels is the accuracy of the labels as they are computed directly on available separated sources audio. Assuming that isolated samples used during scene generation are of good quality in terms of low other source content, the resulting ground truth labels can be considered as reliable. However, the prediction is a regression task which is typically difficult to learn. The best studied masking model relies on third-octave band sound levels, meaning that the learned model needs to perform an implicit source separation. Moreover, once an estimation of the sound levels are obtained the selected masking model must be used as the next step of pleasantness prediction. This model is not fully accurate and thus leads to small errors in pleasantness values even for perfect sound level estimations.\\

Instead, we can choose to directly predict if each of the three sources is present (heard) at a given time. This is either a detection problem if the data is processed asynchronously or a multi-label classification problem if the data is processed in frames, both of which are much simpler tasks. The corresponding labels are directly the outputs of the best studied masking model $\hat T(\alpha, \beta)$. Its uncertainties are thus transferred to "weak" labels rather than considered as part of the constant processing chain used to obtain pleasantness estimations.\\

In this report, two models for source-specific sound level and presence estimation are implemented and compared. \ml{manque un lien entre les deux derniers par}

@@

Two datasets of simulated scenes are created for the proposed task:
\begin{itemize}
\item The development set containing 400 45~s scenes (total 5 hours) is used during the training process
\item The evaluation set containing 200 45~s scenes (total 2.5 hours) is used to compute the model's performance and its generalization capabilities
\end{itemize}

The generation procedure is the same for both datasets, and is equivalent to that of the simulated sub-corpus considered for the perceptual test. Only traffic, human voices and birds sources compose the sound scenes due to the low number of available samples for other sources. The isolated samples database is split in the same proportions as the two datasets: two-thirds for the development set and the remaining one-third for the evaluation set. As a result, while the generation process is the same for both corpora both the scenarios and samples are different.\\

The sensors implemented as part of the Cense project provide fast measurements of sound levels for each third-octave band in the $20~Hz - 12.5~kHz$ range in their normal operation. Thus, the model should take as input the same representation. Third-octave spectra are computed for 125~ms frames and 29 frequency bands of the mixed sound scenes. Inputs are then obtained by splitting the resulting spectrograms into frames with a 1~s duration. This scale is considered relevant for perception and was already used to compute the masking models proposed in the perceptual experiment results analysis. The input frames of dimension 29x8 are processed independently in this study.\\ \ml{cite sensor paper}

@@

Training the models is achieved by progressively minimizing a loss function $L(y, \hat y)$, where $y$ and $\hat y$ represent the target and predicted outputs respectively. For sound level estimation we use the MSE or L2 loss:
\begin{equation}
L(y, \hat y) = \sum_s \left(\hat y_s - y_s\right)^2
\end{equation}
where $s$ is the source, $y_s$ and $\hat y_s$ are the target and predicted sound level of source $s$ in dB. For the prediction of the presence, the cost function is the binary cross-entropy of equation:
\begin{equation}
L(y, \hat y) = -\sum_s y_s log\left(\hat y_s\right) + (1-y_s) log\left(1-\hat y_s\right)
\end{equation}
where $s$ is the source, $y_s$ and $\hat y_s$ are the target and predicted presence for source $s$ in the 0-1 range.\\

@@

Table~\ref{tab:perf_cmp} displays the performance of the two studied models for metrics at the physical level. The first model outputs very poor estimations of the source-specific sound level, but the resulting source presence accuracy is high at 92.51\%. As a result, the time of presence estimates are precise with a root mean-squared error of 0.12 on a 0-1 scale. This behavior may be explained by the fact that $\hat T_s(\alpha)$ indicator is too permissive: if some sources are close in terms of full-band sound level, the threshold $\alpha = -19dB$ allows rougher estimates for binary presence values. Thus, when using $\hat T_s(\alpha)$, it is sufficient to obtain good presence detection that the model output either very low values when the source is absent or poor estimates of the level when it is present. Here, the model fails at the source separation and level estimation task and is probably closer to a detection model.\\ \ml{Clarifier les deux derineire phrases}

The presence detection model performs well for both presence accuracy (92.11\%) and resulting $\hat T_s(\alpha, \beta)$ estimation (0.12 RMSE). Both models obtain lower accuracies for birds than for traffic or voices. However, the origin of errors varies. The sound level model has an very low false positive rate across sources while its false negative rate varies from 7.8\% for traffic to more than 18\% for birds. The opposite is observed for the presence model which displays a higher false positive rate for birds. Overall, traffic is found to be the source that is most easily detected. This may be due to its power spectral density that is concentrated in the lower frequency bands while voice and bird information is usually concentrated in narrower frequency ranges when using third-octave bands.\\

@@

Next, a comparison is made of the models's capacity to result in relevant pleasantness estimates. For this analysis, the perceptual experiment is used as assessments are available. It is further split in two parts:
\begin{enumerate}
\item The 25 recorded and replicated scenes, which contain additional sources not present in the development or evaluation datasets simulated for the optimization of deep learning models,
\item The 75 simulated scenes that are obtained from the same simulation process as both the development and evaluation dataset.
\end{enumerate}
The results are shown in Table~\ref{tab:pppred}. Again, both models lead to similarly correct estimations of pleasantness with errors lower than 1.20 on a 10-point scale. Interestingly, predictions are better on the first tested sub-corpus which contains other sources, indicating that models are quite robust to this type of perturbation in the data. Though, a larger set of scenes and environments with perceptual annotations would be required to further validate this point.\\

@@

Subsequently, a deep learning model with state-of-the-art architecture is developed for this task. The model takes as input the fast third-octave band representation of the mixed scene and is trained on a corpus of simulated scenes to output presence labels in short time frames of 1s. Low pleasantness prediction errors are achieved on the perceptual experiment corpus: around 1.15 root mean-squared error on a 10-point scale which is found satisfying \ml{est que l'on peut comparer a la variance de jugement perceptif ?}.\\

Future work will focus on the study of the robustness of developed models to new environment and data. Here only three sources were considered during training, and a limited isolated sound sources database have been considered. Thus, pleasantness estimation chain needs to be evaluated using data gathered in the experimentation site, where the Cense sensor network is to be implemented. %, as it contains specific sound environments as well as unique sources that were not considered in this study.
